Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-07-21 Document/paper.tex
edit

These two questions have been difficult to address because for a long
time it was difficult to produce realistic galactic discs within the
CDM context.
The success of such enterprise has been the result of high numerical
resolution and the understanding that baryonic effects such as stellar
feedback and black hole feedback play an important role in forming a
galactic disc resembling the Milky Way.
Along the way numerical experiments show that the baryonic effects also impact
the dark matter halo shape making the halo rounder than it otherwise
be in a dark matter only simulation
\citep{Dubinski94,Debattista08,Kazantzidis10,Abadi10,Bryan13,Chua19,Artale19}. 
High resolution simulations have also allowed detailed studies of the
alignment between the stellar disc and the dark matter halo
\citep{Bailin05,DeBuhr12,Debattista13,Gomez17}.

Observational constraints of the dark matter halo have been in
tension during the last decade. 
One can find studies favoring prolate
\citep{Banerjee_and_Chanda_2011,Bowden_et_al._2016},  oblate
\citep{LM10,Deg_and_Widrow_2013,Vera-Ciro_and_Helmi_2013} and
spherical configurations \citep{Bovy16}.  
Constraints from modeling of stellar streams discard the prolate
configuration \citep{LM10,Pearson_et_al._2015,Bovy16} although some other studies
still doubt that stellar streams can be used to constrain the halo
shape once other assumptions, such as the density profile, are relaxed
\citep{Ibata_et_al._2013}.

In this paper we contribute to the debate by reporting measurements of
the DM halo shape in MW type galaxies from state-of-the-art simulations from the
Auriga project.
These simulations have large enough numerical resolution, and explicit
cosmological context and an appropriate feedback physics to produce
realistic MW discs.
This paper is structured as follows. 
In Section \ref{sec:numerical} we present the most relevant details of
the simulations, in Section \ref{sec:method} we present the method
that we use to measure the DM halo shape. 
In Section \ref{sec:results} we present our results focusing on the
radial shape trends at $z=0$ and the alignments with the stellar disc.
We discuss our results in Section \ref{sec:discussion} to place them
into the context of other numerical work, explore correlations of the
shape with baryonic properties in the disc and finally make a direct
comparison against observational constraints for the MW's dark matter halo shape.
We finalize with our conclusions in Section \ref{sec:conclusions}.

>>>>

These two questions have been challening to address because for a long
time it was difficult to produce realistic galactic discs within the
CDM context \citep{1997ApJ...478...13N}.
The success of such enterprise has been the result of high numerical
resolution and the understanding that baryonic effects such as stellar
feedback and black hole feedback play an important role in forming a
galactic disc resembling the Milky Way.
Along the way, many numerical experiments have also show that the baryonic
effects also impact the dark matter halo shape making the halo rounder
than it otherwise be in a dark matter only simulation
\citep{Dubinski94,Debattista08,Kazantzidis10,Abadi10,Bryan13,Chua19,Artale19}. 
But only few simulations have presented detailed results on the
radial evolution of the alignment between the stellar disc and the
dark matter halo \citep{Bailin05,Debattista13}.

Observational constraints of the dark matter halo have been in
tension and rapidly evolving during the last decade. 
One can find studies favoring prolate
\citep{Banerjee_and_Chanda_2011,Bowden_et_al._2016},  oblate
\citep{LM10,Deg_and_Widrow_2013,Vera-Ciro_and_Helmi_2013} and
spherical configurations \citep{Bovy16}.  
Constraints from modeling of stellar streams discard the prolate
configuration \citep{LM10,Pearson_et_al._2015,Bovy16} although some other studies
still doubt that stellar streams can be used to constrain the halo
shape once other assumptions, such as the density profile, are relaxed
\citep{Ibata_et_al._2013}.

In this paper we analyze the internal DM halo shape in thirty MW type
galaxies from state-of-the-art simulations from the Auriga project.
These simulations have large enough numerical resolution, and explicit
cosmological context and an appropriate feedback physics to produce
realistic MW discs.
Futhermore, the relatively large number of simulated systems allows us
to have a handle on the statistical significance of our results.
This paper is structured as follows. 
In Section \ref{sec:numerical} we present the most relevant details of
the simulations, in Section \ref{sec:method} we present the method
that we use to measure the DM halo shape. 
In Section \ref{sec:results} we present our results focusing on the
radial shape trends at $z=0$ and the alignments with the stellar disc.
We discuss our results in Section \ref{sec:discussion} to place them
into the context of other numerical work, explore correlations of the
shape with baryonic properties in the disc and finally make a direct
comparison against observational constraints for the MW's dark matter halo shape.
We finalize with our conclusions in Section \ref{sec:conclusions}.
